Laves polyhedrons in synthetic tennantite, Cu\textsubscript{12}As\textsubscript{4}S\textsubscript{13}, and the lattice dynamics

### 2021-01-17 paper_tennantite.tex

\subsection{Single-crystal X-ray diffraction}\label{sec:level2}
A spherical shape single-crystal sample of Cu\textsubscript{12}As\textsubscript{4}S\textsubscript{13}  was prepared for the anisotropic extinction study.
Enraf-Nonius CAD-4 diffractometer with AgK\textsubscript{$\alpha$} and graphite monochromator was used.
The spherical shape single-crystal structural study results are presented in Appendix, they are compared with the structural data of the irregular form sample\cite{yaroslavzev2019}.

@@

\subsection{Remarks on the structure of synthetic tennantite}\label{sec:level2}
Study of the high-grade single crystal allowed us to re-investigate the crystal structure of Cu\textsubscript{12}As\textsubscript{4}S\textsubscript{13}  with a high precision.
Our data confirms that the compound contains 12 copper atoms per formula unit, and shows no evidences of a copper deficiency.
Neither XRD, nor TEM data shown defects.
Due to the crystal growth method used, we suspected a possibility of anisotropic extinction.
The intensity vs  $\psi$ angle were measured for several reflections to analyze this effect.

The results are presented in Fig.~\ref{fig:rock_curv} .
All curves were approximated by the least-square fitting.
Only for \hkl[4 0 0], \hkl[0 4 0], and \hkl[0 0 4] reflections the coefficients with $x$ and $x^{2}$ were in the range of 5--8$\sigma$, while for the two other sets they were about $\sigma$ or less (with one 2$\sigma$ exception).
Thus, we conclude that the intensities of weak reflections at small angles and reflections at high angles have fixed $\psi$ angles.
On the contrary, $\psi$ angles of strong reflections with small $\sin\theta/\lambda$ vary, that may indicate an anisotropic extinction.

\begin{figure*}
\centering
\subfigure{\includegraphics[width=0.3\textwidth]{200}}
 \quad
\subfigure{\includegraphics[width=0.3\textwidth]{400}}
 \quad
\subfigure{\includegraphics[width=0.3\textwidth]{1200}}
\caption{\label{fig:rock_curv} Normalized intensities of reflections vs. $\psi$ angle; left --- 200, 020, 002 (weak, sin=0.098), middle --- 400, 040, 004 (strong, sin=0.197), right --- 1200, 0120, 0012 (moderate, sin=0.590).}
\end{figure*}

@@

The extinction can be explained by the dynamic disorder of Cu21 atoms. Such a disorder is observed in the analyzed HAADF image of the sample in [110] zone (Fig.~\ref{fig:micro}).
The electron diffraction, and overlay of HAADF image with the tennantite crystal [110] zone generated by VESTA\cite{Momma2011}, are shown in Appendix.
The processed HAADF image (Fig.~\ref{fig:micro} right) shows  the ellipticity for the Cu/As atomic columns, which is a measure of the atomic column elongation\cite{Nord2017}.
The atomic column elongation is explained by the blurring of the intensity distribution, and a large atomic displacement.
A similar inhomogeneous electron density was observed in tetrahedrite\cite{Mishra2017}, and tetrahedral framework structure\cite{Suekuni2019}.
Also, the anisotropy may be caused by the crystal growth conditions or/and a crystal cell distortion.
If so, then the structure would have a diagonal distortion or/and have more or less than 12 copper atoms per cell, which is not the case.

>>>>

\subsection{Single-crystal X-ray diffraction}\label{sec:level2}
A spherical shape single-crystal sample of Cu\textsubscript{12}As\textsubscript{4}S\textsubscript{13}  was prepared for the anisotropic extinction study\cite{Becker1975}.
Enraf-Nonius CAD-4 diffractometer with AgK\textsubscript{$\alpha$} and graphite monochromator was used. 
The spherical shape single-crystal structural study results are presented in Appendix, they are compared with the structural data of the irregular form sample\cite{yaroslavzev2019}.

@@

\subsection{Remarks on the structure of synthetic tennantite}\label{sec:level2}
Study of the high-grade single crystal allowed us to re-investigate the crystal structure of Cu\textsubscript{12}As\textsubscript{4}S\textsubscript{13}  with a high precision.
Our data confirms that the compound contains 12 copper atoms per formula unit, and shows no evidences of a copper deficiency.
Neither XRD, nor TEM data shown defects.

@@

Due to the crystal growth method used, anisotropic extinction in the sample takes place. Extinction parameters are presented in the Table~\ref{tab:anis_ext}.
Also the extinction can be explained by the dynamic disorder of Cu21 atoms. Such a disorder is observed in the analyzed HAADF image of the sample in [110] zone (Fig.~\ref{fig:micro}).
The electron diffraction, and overlay of HAADF image with the tennantite crystal [110] zone generated by VESTA\cite{Momma2011}, are shown in Appendix.
The processed HAADF image (Fig.~\ref{fig:micro} right) shows  the ellipticity for the Cu/As atomic columns, which is a measure of the atomic column elongation\cite{Nord2017}.
The atomic column elongation is explained by the blurring of the intensity distribution, and a large atomic displacement.
A similar inhomogeneous electron density was observed in tetrahedrite\cite{Mishra2017}, and tetrahedral framework structure\cite{Suekuni2019}.
Also, the anisotropy may be caused by the crystal growth conditions or/and a crystal cell distortion.
If so, then the structure would have a diagonal distortion or/and have more or less than 12 copper atoms per cell, which is not the case.
